Studying public transit API query logs\\ to get an indication of travel flows

### 2015-12-21 ldfaxis2.tex
rewrote axises, rewrote abstract, added chord diagrams, etc.

    \draw[axis] (-100, 0) -- (0, 0);
    \node[axislabel,align=left,anchor=north west] at (-100, -2.3)
    {
%      generic requests\\
      high client effort\\
      high server availability\\
      know how many times datadump was downloaded
    };
    \node[axislabel,align=right,anchor=north east] at (0, -2.3)
    {
%      specific requests\\
      high server effort\\
      low server availability\\
      detailled query logs
    };

    \tick{-93}{GTFS data\\dump}
    \tick{-10}{route planning\\service}
    \tick{-50}{Linked\\Connections}

    \vaguetick{-84}    \vaguetick{-70}
    \vaguetick{-10}
    \vaguetick{-28}
    \vaguetick{-20}

   %\node[align=center,font=\fontsize{8}{9}\selectfont\bfseries] at (0, -8.3)
   %   {various types of\\Linked Data Fragments};
  \end{tikzpicture}
  \caption{This axis extends \cref{fig:LDFAxis1} with Linked Connections: a different choice of trade-offs for route planning advice over the Web.
  }
  \label{fig:LDFAxis2}
\end{figure}

>>>>

    \draw[axis] (-100, 0) -- (-5, 0);
    \node[axislabel,align=left,anchor=north west] at (-100, -2.3)
    {
%      generic requests\\
      high client effort\\
      high server availability\\
      know how many times datadump was downloaded
    };
    \node[axislabel,align=right,anchor=north east] at (-5, -2.3)
    {
%      specific requests\\
      high server effort\\
      low server availability\\
      detailled query logs
    };
    \node[text,ticklabel] at (-53, 2.4) {Linked\\Connections};
    \tick{-90}{GTFS data\\dump}
    \tick{-15}{route planning\\service}
    \vaguetick{-23}
    \vaguetick{-32}
    \vaguetick{-41}
    \vaguetick{-50}
    \vaguetick{-60}
    \vaguetick{-70}
    \vaguetick{-80}

   %\node[align=center,font=\fontsize{8}{9}\selectfont\bfseries] at (0, -8.3)
   %   {various types of\\Linked Data Fragments};
  \end{tikzpicture}
  \caption{This axis extends \cref{fig:LDFAxis1} with Linked Connections: the different ticks on the axis for client effort versus server expressivity.}
  \label{fig:LDFAxis2}
\end{figure}

### 2015-12-21 paper.tex
rewrote axises, rewrote abstract, added chord diagrams, etc.

\title{What we can learn from one year of\\ public transit route planning API queries}
\numberofauthors{6}
\author{
\alignauthor
Pieter Colpaert\\
\affaddr{\email{\texttt{pieter.colpaert@ugent.be}}}
\and
\alignauthor
Alvin Chua\\
\affaddr{\email{\texttt{alvin.chua@asro.kuleuven.be}}}
\and
\alignauthor
Ruben Verborgh\\
\affaddr{\email{\texttt{ruben.verborg@ugent.be}}}
\and
\alignauthor
Erik Mannens\\
\affaddr{\email{\texttt{erik.mannens@ugent.be}}}
\and
\alignauthor
Rik Van de Walle\\
\affaddr{\email{\texttt{rik.vandewalle@ugent.be}}}
\and
\alignauthor
Andrew Vande Moere\\
\affaddr{\email{\texttt{andrew.vandemoere\\@asro.kuleuven.be}}}\\
}

\maketitle
\begin{abstract}
% Context
%Public transit schedules are made available on the Web in various ways, such as in a GTFS file, a route planning API, or through Linked Connections.
The iRail project hosts a route planning API for the Belgian railway company.
This webservice provides a concise answer to a complex question with parameters such as ``departuretime'', ``from'' and ``to''.
% Need
In the field of urban planning, researchers need an indication of how people move between cities. 
Yet, getting this data from official sources has proven to be troublesome.
% Task
In this paper, we analyse the query logs from the iRail API, to tell us something about how people move between cities in Belgium.
% Object
We studied queries for the year 2013, which contains an average of 3000 queries per day.
% Findings
We found X, Y, Z interesting patterns (\todo{Alvin}), which correspond with reality.
% Conclusion
This proves that query logs from route planning systems are valuable as an indication of transit flows and may house interesting stories.
% Perspectives
However, many private owned route planners exists, which keep these query logs strictly confidential.
Bearing this in mind, we discuss the current state of the art in public transit schedule publishing, and formulate opportunities for gathering query logs when using a Linked Data Fragments approach.

@@

\section{Results}
\label{sec:results}
\todo{
We can also observe the effect of "density" in Flanders where people move from the rural towards cities. Look at weekday-mornings: commuting towards Gent, Brussels, Antwerp. This is not as obvious in the Walloon region (except for Liege). The reverse occurs on weekday-evenings. Local patterns can also be observed and this corresponds to what we know: "cities are the central hubs".
Walloon region: everything converges on Brussels
I found that whereas there's almost no movement from Antwerp city in the week to West Flanders, there is movement from Antwerp in the weekend to the coast and back. 
}

@@

\todo{add a couple of nice images with explanation here}

@@

\section{Publishing transport data and the impact on query logs}
\label{sec:publishing}

Publishing public transport data today happens in two ways: by publishing a data dump of the timetables or by hosting a public route planning service.
Linked Data Fragments~\cite{ldf} is a way to think about data publishing as a set of trade-offs between client and servers.
These trade-offs can be mapped on an axis, cfr. \cref{fig:LDFAxis2}, in which the 2 extremes are identified as: the server does all the processing versus the client does all the processing.

Also for query logs we can make the same trade-off, in which the 2 extremes are the server knows only when someone downloads the entire datadump versus the server has a detailled query log.
In this paper, we have studied the query logs of an API s

@@

>>>>

\title{Studying public transit API query logs\\ to get an indication of travel flows}
\numberofauthors{6}
\author{
\alignauthor
Pieter Colpaert\\
\affaddr{\email{\texttt{pieter.colpaert@ugent.be}}}
\and
\alignauthor
Alvin Chua\\
\affaddr{\email{\texttt{alvin.chua@asro.kuleuven.be}}}
\and
\alignauthor
Ruben Verborgh\\
\affaddr{\email{\texttt{ruben.verborg@ugent.be}}}
\and
\alignauthor
Erik Mannens\\
\affaddr{\email{\texttt{erik.mannens@ugent.be}}}
\and
\alignauthor
Rik Van de Walle\\
\affaddr{\email{\texttt{rik.vandewalle@ugent.be}}}
\and
\alignauthor
Andrew Vande Moere\\
\affaddr{\email{\texttt{andrew.vandemoere\\@asro.kuleuven.be}}}\\
}

\maketitle
\begin{abstract}
% Context
%Public transit schedules are made available on the Web in various ways, such as in a GTFS file, a route planning API, or through Linked Connections.
The iRail project hosts a route planning API for the Belgian railway company.
This webservice provides an answer to a route planning question with parameters such as ``departuretime'', ``from'' and ``to''.
% Need
In the field of urban planning, researchers need an indication of how people move between cities. 
%Yet, getting this data from official sources has proven to be troublesome.
% Task
In this paper, we analyse the query logs of the iRail API, to see whether it can tell us something about travel flows between cities in Belgium.
% Object
We studied queries for the year 2013, which contains an average of $\sim3000$ interesting queries per day ($\sim0.4\%$ of the amount of passengers per weekday).
% Findings
We illustrate \todo{X} interesting patterns that correspond to reality and we show an outlier in the data.
% Conclusion
Visualizing the data shows the query logs are a better representation than random data, yet more research is needed to understand how representative for travel flows they may be.
% Perspectives
In earlier work, we have suggested the Linked Data Fragments axis to discuss new approaches to publishing queryable public transit data for maximized use.
We suggest a discussion for gathering query logs when using Linked Data Fragments approaches.

@@

\section{Results}
\label{sec:results}

\todo{
We can also observe the effect of "density" in Flanders where people move from the rural towards cities. Look at weekday-mornings: commuting towards Gent, Brussels, Antwerp. This is not as obvious in the Walloon region (except for Liege). The reverse occurs on weekday-evenings. Local patterns can also be observed and this corresponds to what we know: "cities are the central hubs".
Walloon region: everything converges on Brussels
I found that whereas there's almost no movement from Antwerp city in the week to West Flanders, there is movement from Antwerp in the weekend to the coast and back. 
}

\begin{figure}
\centering
\includegraphics[width=8.1cm,align=center]{brussels}
\caption{Double chord visualization of the city of Brussels: transit flows towards Brussels from elsewhere in Belgium and vice versa. Brussels acts as the central hub in the star-shaped railway network of Belgium.}
\label{fig:brussels}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=8.1cm,align=center]{antwerp}
\caption{Double chord visualization of the city of Antwerp: transit flows towards Antwerp from elsewhere in Belgium and vice versa. The biggest transit flows to elsewhere in the province of Antwerp Central station is the hub for the province.}
\label{fig:antwerp}
\end{figure}

@@

\subsection{Pattern 3: cannot be true}
\label{sec:pattern3}

\section{Publishing transport data: the impact on query logs}
\label{sec:publishing}

Publishing public transport data today happens in two ways: by publishing a data dump of the timetables or by hosting a public route planning service.
Linked Data Fragments~\cite{ldf} is a way to think about data publishing as a set of trade-offs between client and servers.
These trade-offs can be mapped on an axis, cfr. \cref{fig:LDFAxis2}, in which the 2 extremes are identified as: the client does all the processing by downloading all the data at once, versus the server does all the processing by exposing \emph{expressive} APIs.
Linked Connections~\cite{lc} proposes other ticks on this axis.
By publishing fragments of the data needed for by the shortest path algorithm, the algorithm can be executed on the client.
Furthermore, it lowers to load on the server, and it enables clients to federate queries over different servers, including different regions, transit modes and user's need.

@@

The expressivity of the server also affects the way logs can be gathered.
When the server becomes less expressive, e.g., when it only allows to set a departure station and a departure time, then the server doesn't know the arrival station, which makes it harder to get an indication of the travel flow.
